Prediction of functional markers of mass cytometry data via deep learning

### 2019-01-24 revision/latex_template/author.tex
added revision files

\documentclass[conference]{IEEEtran}

%========================================================================= PACKAGES
\usepackage{amsmath,amssymb}
\usepackage[mathscr]{euscript} %To be able to use script letters with \mathscr{}
\usepackage{amsthm} %To be able to write Definitions, Theorems, etc.
\usepackage{tikz} %To draw graphs.
\usepackage{graphicx} % To scale equations
\newcommand*{\Scale}[2][4]{\scalebox{#1}{\ensuremath{#2}}} % To scale equations
\usepackage{framed} % To be able to frame theorems and stuff
%\usepackage{float} %To be able to place figures exactly where I want with [H]
\usepackage{multirow}
\usepackage{color}
\usepackage{subfigure}
\usepackage{longtable}
\usepackage{MnSymbol}
\usepackage{xcolor}  \definecolor{shadecolor}{rgb}{.95,.95,.95}  %To put a shaded region
\usepackage[numbers,square, comma, sort&compress]{natbib}	 %Make the references [1-8], rather than [1]-[8], as with \usepackage{cite}
\renewcommand*{\bibfont}{\footnotesize} %Size of the references; only needed with natbib
%\usepackage{makeidx} %To create an index
%\makeindex
\usepackage[hidelinks]{hyperref} %To be able to use links inside the document; [hidelinks removes the ugly red boxes]

%==================================================== ALGORITHM SETUP
\usepackage[ruled,vlined]{algorithm2e}

%========================================================================= THEOREMS, LEMMAS, DEFINITIONS, ETC.
\newtheorem{myDefinition}{Definition}
\newtheorem{myTheorem}{Theorem}
\newtheorem{myLemma}{Lemma}
\newtheorem{myCorollary}{Corollary}
\newtheorem{myProposition}{Proposition}
\newtheorem{myExample}{Example}
\newtheorem{myRemark}{Remark}
\newtheorem{myConjecture}{Conjecture}
\newtheorem{myQuestion}{Question}
\newtheorem{myProblem}{Problem}

%========================================================================= STUFF TO DRAW GRAPHS
\tikzstyle{every edge}=  [draw]
\tikzstyle{vertex} = [draw,circle,minimum size=1pt]
\tikzstyle{label} = [minimum size=.1pt,font=\scriptsize]
\tikzstyle{title} = [minimum size=.25cm,font=\small]

%========================================================================= SYMBOLS
%\newcommand{\var}[1]{\mathsf{#1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\careful}{\textcolor{red}}
\newcommand{\comment}{\textcolor{blue}}

%Miscelaneous
%\def \R{\mathbb{R}}
%\def \P{\mathsf{P}}
%\def \T{\mathsf{T}}
%\def \c{\mathsf{c}}
\def \spn{{\rm span}}
\def \rnk{{\rm rank}}

%Scalars
\def \r{{\hyperref[rDef]{r}}}

%text
\def \as{{\hyperref[asDef]{{\rm a.s.}}}}
%Assumptions
\def \Cone{{\hyperref[ConeDef]{{\rm {\bf C1}}}}}
%Theorems
\def \mainThm{{\hyperref[mainThm]{Theorem \ref{mainThm}}}}
%Sections
\def \modelSec{{\hyperref[modelSec]{Section \ref{modelSec}}}}
%Figures
\def \distributionFig{{\hyperref[distributionFig]{Figure \ref{distributionFig}}}}
%Algorithms
\def \validationAlg{{\hyperref[validationAlg]{Algorithm \ref{validationAlg}}}}

\newcounter{@inst}
\newcounter{@auth}
\newcounter{auco}
\newdimen\instindent
\newbox\authrun
\newtoks\authorrunning
\newtoks\tocauthor
\newbox\titrun
\newtoks\titlerunning
\newtoks\toctitle

\def\clearheadinfo{\gdef\@author{No Author Given}%
                   \gdef\@title{No Title Given}%
                   \gdef\@subtitle{}%
                   \gdef\@institute{}%
                   \gdef\@thanks{}%
                   \global\titlerunning={}\global\authorrunning={}%
                   \global\toctitle={}\global\tocauthor={}}

\def\institute#1{\gdef\@institute{#1}}

\def\institutename
 {\begingroup
 \if!\@institute!\else
 \def\thanks##1{\unskip{}}%
 \def\iand{\\[5pt]\let\and=\nand}%
 \def\nand{\ifhmode\unskip\nobreak\fi\ $\cdot$ }%
 \let\and=\nand
 \def\at{\\\let\and=\iand}%
 \footnotetext[0]{\kern-\bibindent
 \ignorespaces\@institute}\vspace{5dd}\fi
 \endgroup
 }%

% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor}

@@

\title{Prediction of functional markers of mass cytometry data via
  deep learning}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
%\thanks{Both authors contributed equally to this manuscript}
\author{Claudia Sol\'is-Lemus, Xin Ma, Maxwell Hostetter II, Suprateek
  Kundu, Peng Qiu, Daniel Pimentel-Alarc\'on}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Claudia Sol\'is-Lemus \at Emory University, Atlanta, GA, \email{csolisl@emory.edu}
\and Xin Ma \at Emory University, Atlanta, GA,
\email{xin.ma@emory.edu}
\and Maxwell Hostetter II \at Georgia State University, Atlanta, GA,
\email{mhostetter1@student.gsu.edu}
\and Suprateek Kundu \at Emory University, Atlanta, GA,
\email{suprateek.kundu@emory.edu}
\and Peng Qiu \at Georgia Institute of Technology and Emory University, Atlanta, GA,
\email{peng.qiu@bme.gatech.edu}
\and Daniel Pimentel-Alarc\'on \at Georgia State University, Atlanta, GA,
\email{pimentel@gsu.edu}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\begin{abstract}
Recently, there has been an increasing interest in the analysis of
flow cytometry data, which involves measurements of a set of surface
and functional markers across hundreds and thousands of cells. These
measurements can often be used to differentiate various cell types and
there has been a rapid development of analytic approaches for
achieving this. However, in spite of the fact that measurements are
available on such a large number of cells, there have been very
limited advances in deep learning approaches for the analysis of flow
cytometry data. Some preliminary work has focused on using deep
learning techniques to classify cell types based on the cell protein
measurements. In a first of its' kind study, we propose a novel deep
learning architecture for predicting functional markers in the cells
given data on surface markers. Such an approach is expected to
automate the measurement of functional markers across cell samples,
provided data on the surface markers are available, that has important
practical advantages. We validate and compare our approach with
competing machine learning methods using a real flow cytometry
dataset, and showcase the improved prediction performance of the deep
learning architecture.
\end{abstract}

@@

\begin{figure*}
\centering
\includegraphics[scale=0.35]{figures/surface-hist.pdf}
\includegraphics[scale=0.35]{figures/functional-hist.pdf}
\includegraphics[scale=0.35]{figures/scatterplot.pdf}
\caption{Exploratory plots of surface and functional markers. The
  histograms show a biased pattern, and the scatterplot shows non
  linearity, both violations of crucial assumptions in standard
  regression models.}
\label{exploratory}
\end{figure*}

@@

@@

\begin{figure*}
\centering
\includegraphics[scale=0.4]{figures/neuralNet.eps}
\caption{Neural network for predicting functional markers (18
  responses) from surface markers (15 predictors) with 4 hidden
  layers with 90, 90, 45 and 45 nodes each.}
\label{nn}
\end{figure*}

@@

@@

\begin{figure}
\centering
\includegraphics[scale=0.45]{figures/vector-mse.pdf}
\includegraphics[scale=0.45]{figures/time.pdf}
\caption{Left: Vector MSE (equation \ref{vectorMSEformula}) for all
  four methods sorted from most accurate (neural network) to least
  accurate (decision tree). Right: Running time (in seconds) for the
  training and validation sets (sample 750,000 rows) for all four
  methods, sorted from fastest (linear regression) to slowest (random
  forest).}
\label{vectorMSE}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[scale=0.35]{figures/violin.pdf} \hspace{0.25cm}
\includegraphics[scale=0.35]{figures/response1.pdf}
\hspace{0.25cm}
\includegraphics[scale=0.35]{figures/response14.pdf}
\caption{Left: Violin plot for four functional markers (responses). Horizontal lines represent the 25th quantile, median and 75th quantile. Center: Predicted vs observed responses on the first functional marker (141.pPLCgamma2)
  across all four methods. Left: Predicted vs observed responses on the 14th functional marker (171.pBtk.Itk)
  across all four methods. The closer the slope to 1 (black line), the
  better.}
\label{violin}
\end{figure*}

@@

\begin{figure}[ht]
\centering
\includegraphics[scale=0.3]{figures/scatterplot1.pdf} \hspace{0.5cm}
\includegraphics[scale=0.3]{figures/scatterplot14.pdf}
\caption{Scatterplot of selected surface markers (predictors) and selected functional markers (responses). Left: the first response (functional marker 141.pPLCgamma2) shows a linear relationship to the predictor (surface marker 115.CD45), which partially explains the better MSE in figure \ref{MSE}. Right: the 14th response (functional marker 171.pBtk.Itk) shows a non-linear relationship to the predictor (surface marker 115.CD45), which partially explains the worse MSE in figure \ref{MSE}.}
\label{predObs}
\end{figure}

@@

% %======================================================= MAIN THEOREM
% \begin{framed}
% \begin{myTheorem}
% \label{mainThm}
% Write Theorem
% \end{myTheorem}
% \end{framed}

>>>>

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% track changes
\usepackage[normalem]{ulem}
\usepackage[dvipsnames]{xcolor}
\newcommand{\claudia}[2]{\sout{#1}\textcolor{magenta}{\textbf{#2}}}
\newcommand{\missing}[1]{\textcolor{red}{\textbf{#1}}}

@@

\title*{Prediction of functional markers of mass cytometry data via
  deep learning}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
%\thanks{Both authors contributed equally to this manuscript}
\author{Claudia Sol\'is-Lemus, Xin Ma, Maxwell Hostetter II, Suprateek
  Kundu, Peng Qiu, Daniel Pimentel-Alarc\'on}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Claudia Sol\'is-Lemus \at Emory University, Atlanta, GA, \email{csolisl@emory.edu}
\and Xin Ma \at Emory University, Atlanta, GA,
\email{xin.ma@emory.edu}
\and Maxwell Hostetter II \at Georgia State University, Atlanta, GA,
\email{mhostetter1@student.gsu.edu}
\and Suprateek Kundu \at Emory University, Atlanta, GA,
\email{suprateek.kundu@emory.edu}
\and Peng Qiu \at Georgia Institute of Technology and Emory University, Atlanta, GA,
\email{peng.qiu@bme.gatech.edu}
\and Daniel Pimentel-Alarc\'on \at Georgia State University, Atlanta, GA,
\email{pimentel@gsu.edu}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract{Recently, there has been an increasing interest in the analysis of flow cytometry data, which involves measurements of a set of surface and functional markers across hundreds and thousands of cells. These measurements can often be used to differentiate various cell types and there has been a rapid development of analytic approaches for achieving this. However, in spite of the fact that measurements are available on such a large number of cells, there have been very limited advances in deep learning approaches for the analysis of flow cytometry data. Some preliminary work has focused on using deep learning techniques to classify cell types based on the cell protein measurements. In a first of its' kind study, we propose a novel deep learning architecture for predicting functional markers in the cells given data on surface markers. Such an approach is expected to automate the measurement of functional markers across cell samples, provided data on the surface markers are available, that has important practical advantages. We validate and compare our approach with competing machine learning methods using a real flow cytometry dataset, and showcase the improved prediction performance of the deep learning architecture.}

@@

@@

\begin{figure}
\includegraphics[scale=0.3]{figures/surface-hist.pdf}
\includegraphics[scale=0.3]{figures/functional-hist.pdf}
\includegraphics[scale=0.3]{figures/scatterplot.pdf}
\caption{Exploratory plots of surface and functional markers. The
  histograms show a biased pattern, and the scatterplot shows non
  linearity, both violations of crucial assumptions in standard
  regression models.}
\label{exploratory}
\end{figure}

@@

@@

\begin{figure}
\centering
\includegraphics[scale=0.4]{figures/neuralNet.eps}
\caption{Neural network for predicting functional markers (18
  responses) from surface markers (15 predictors) with 4 hidden
  layers with 90, 90, 45 and 45 nodes each.}
\label{nn}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.35]{figures/vector-mse.pdf}
\includegraphics[scale=0.35]{figures/time.pdf}
\caption{Left: Vector MSE (equation \ref{vectorMSEformula}) for all
  four methods sorted from most accurate (neural network) to least
  accurate (decision tree). Right: Running time (in seconds) for the
  training and validation sets (sample 750,000 rows) for all four
  methods, sorted from fastest (linear regression) to slowest (random
  forest).}
\label{vectorMSE}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.3]{figures/violin.pdf} \hspace{0.25cm}
\includegraphics[scale=0.3]{figures/response1.pdf}
\hspace{0.25cm}
\includegraphics[scale=0.3]{figures/response14.pdf}
\caption{Left: Violin plot for four functional markers (responses). Horizontal lines represent the 25th quantile, median and 75th quantile. Center: Predicted vs observed responses on the first functional marker (141.pPLCgamma2)
  across all four methods. Left: Predicted vs observed responses on the 14th functional marker (171.pBtk.Itk)
  across all four methods. The closer the slope to 1 (black line), the
  better.}
\label{violin}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.3]{figures/scatterplot1.pdf} \hspace{0.5cm}
\includegraphics[scale=0.3]{figures/scatterplot14.pdf}
\caption{Scatterplot of selected surface markers (predictors) and selected functional markers (responses). Left: the first response (functional marker 141.pPLCgamma2) shows a linear relationship to the predictor (surface marker 115.CD45), which partially explains the better MSE in figure \ref{MSE}. Right: the 14th response (functional marker 171.pBtk.Itk) shows a non-linear relationship to the predictor (surface marker 115.CD45), which partially explains the worse MSE in figure \ref{MSE}.}
\label{predObs}
\end{figure}

@@
